White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2017-09-12 papers/wdmerger_II/paper.tex
Update timestepping section with new data

For the standard self-heating burns that we use, the most important parameters we
examined come from the timestep limiting scheme described in \autoref{sec:hydrocoupling}.
The parameters $f_{be}$ (from \autoref{eq:timestep_e}) and $f_{bX}$ (from
\autoref{eq:timestep_X}) control the size of the timestep as a function of the
burning rate. As the timestep gets smaller, the error due to the Strang splitting
scheme coupling the reactions and the burning also gets smaller, and we expect
that the results become more accurate.
\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
  \caption{Nickel production as a function of the timestep factors $f_{be}$
           (solid blue, limiting the timestep on changes in internal energy) and
           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
           the timestep by at least an order of magnitude more.
           \label{fig:timestep_factor}}
\end{figure}
\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
production. As either $f$ becomes small, the restriction on the timestep becomes quite
severe: for $f_{bX} = 0.01$, the minimum timestep in the simulation is smaller than
$2 \times 10^{-9}\ \text{s}$ and the full simulation requires over 750,000 timesteps.
Yet it is clear that even for this steep value of the limiter, which is much stricter
than the limiter used in, e.g., \cite{hawley:2012} and \cite{raskin:2010}, the
simulation has not fully converged in nickel production. These limiters are somewhat
redundant in the sense that a given $f_{bX}$ corresponds to another (higher) $f_{be}$,
as is evident by the fact that on the graph the curves look similar in shape but with
a horizontal offset.

We also examined the effect of the method used to enforce the limiting, using the
four limiting modes discussed in \autoref{sec:hydrocoupling}. The results are shown
in \autoref{table:burninglimitermode} for the case $f_{be} = 0.3$ (we did not use
limiting with $f_{bX}$ for this test). The two modes that use the full last timestep's
worth of information to estimate $\dot{e}$ yield a slightly larger nickel mass than the
two modes that only use more recent information, but the difference is negligible.

\input{plots/burning_limiter_mode_m_P_0.64_m_S_0.64.tbl}

@@

For all remaining tests in this section, unless otherwise stated, we take $f_{be} = 10^2$
and disable limiting with $f_{bX}$ by setting it to a large number. This choice is a balance
between accuracy and efficiency: we avoid the obviously wrong answers that occur for $f_{be} > 10^3$
while ensuring that the timesteps are long enough that the following tests are computationally
inexpensive.

>>>>

For the standard self-heating burns that we use, the most important parameters we
examined come from the timestep limiting scheme described in \autoref{sec:hydrocoupling}.
The parameters $f_{be}$ (from \autoref{eq:timestep_e}) and $f_{bX}$ (from
\autoref{eq:timestep_X}) control the size of the timestep as a function of the
burning rate. As the timestep gets smaller, the error due to the Strang splitting
scheme coupling the reactions and the hydro also gets smaller, and we expect
that the results become more accurate.
\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
  \caption{Nickel production as a function of the timestep factors $f_{be}$
           (solid blue, limiting the timestep on changes in internal energy) and
           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
           the timestep by at least an order of magnitude more.
           \label{fig:timestep_factor}}
\end{figure}
\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
production (these limiters are somewhat redundant in the sense that a given $f_{bX}$
roughly corresponds to another (higher) $f_{be}$, as is evident by the fact that on
the graph the curves look qualitatively similar in shape but with a horizontal offset.
As either $f$ becomes small, the restriction on the timestep becomes quite severe:
for $f_{be} = 0.001$, the minimum timestep in the simulation is smaller than
$5 \times 10^{-9}\ \text{s}$ and the full simulation requires over 100,000 timesteps.
Such a simulation would be prohibitively expensive for a high resolution 3D calculation.
Yet it is clear that even for this steep value of the limiter, which is much stricter
than the limiter used in, e.g., \cite{hawley:2012} and \cite{raskin:2010}, the
simulation has not fully converged in nickel production.

We also examined the effect of the method used to enforce the limiting, using the
four limiting modes discussed in \autoref{sec:hydrocoupling}. The two modes that
use the full last timestep's worth of information to estimate $\dot{e}$ (Mode 3 and Mode 4)
yield a slightly larger nickel mass than the two modes that only use more recent information,
but the difference is less than 10\%, and it is not clear which is preferable since
for the value of the timestep limiter we used to compare these modes ($f_{be} = 0.3$),
the simulation is not converged anyway.

@@

For all remaining tests in this section, unless otherwise stated, we take $f_{be} = 10^2$
and disable limiting with $f_{bX}$ by setting it to a large number. This choice was made
somewhat arbitrarily for computational efficiency.
